For Finitary Induction-Induction, \\ Induction is Enough

### 2020-04-29 paper/paper.tex
stuff

\begin{example}[Example signatures]\label{ex:signatures}
  The signature for natural numbers is given by the function
  \[
    \lambda M.(\boldsymbol{\cdot}^M\,\ext^M\,\U^M\,\ext^M\,\El^M\,0^M\,\ext^M\,1^M \Ra^M \El^M\,1^M).
  \]
  For readability we write the same signature without the $\lambda$
  and the superscripts below on the left and we also write an informal
  version with variable names on the right for comparison.
  \[
  \boldsymbol{\cdot}\,\ext\,\U\,\ext\,\El\,0\,\ext\,1 \Ra \El\,1 \hspace{5.9em}
  \boldsymbol{\cdot}\,\ext\,N:\U\,\ext\,z : \El\,N\,\ext\,s : N \Ra \El\,N
  \]
  This signature is given by a context of length three. We start with
  the empty context $\boldsymbol{\cdot}$, then we declare a sort by
  $\U$, then we declare an operator producing an element of $\El\,0$
  where $0$ is a de Bruijn index referring to the sort $\U$. Finally
  we declare an operator which takes as input an element of the sort
  $1$ and produces an element of the same sort. Note the assymetry of
  the function type $\Ra$: on the left there needs to be an element of
  $\U$, while on the right any type can appear (including another
  function type). This ensures strict positivity of the operators.

@@

  $\Con$, $\Ty$, $\Sub$, and so on, one per line. Each such components

@@

\begin{definition}[A model of ETT supports IITs]
  A model of ETT supports IITs if for any signature signature
  $\Omega:(M:\SignAlg)\ra\Con^M$ we have an $$
\end{definition}

Now, given a signature $f:(M:\SignAlg)\ra\Con^M$, we obtain
$(f\,\ADS)^\A$

  We describe the most important
  components, a full definition is given in the appendix (TODO). A
  $\Gamma:\Con$ is given by a $\Sigma$ type with three elements:
  \[
    (\Gamma^\A:\Set) \times (\Gamma^\D:\Gamma^\A\ra\Set) \times (\Gamma^\S:(\gamma:\Gamma^\A)\ra\Gamma^\D\,\gamma\ra\Set)
  \]
  As a signature is a context in an arbitrary signature algebra, using
  $\ADS$ a signature provides an algebra $\Gamma^\A$, a displayed
  algebra $\Gamma^\D$ and a section $\Gamma^\S$.

  A type in $\ADS$ is given by dependent versions of $\Con$:
  \begin{alignat*}{6}
    & (A:\Ty\,\Gamma) := \,\,&& (A^\A:\Gamma^A\ra\Set) \times (A^\D:\Gamma^\D\,\gamma\ra A^\A\,\gamma\ra\Set) \times \\
    & && (A^\S:\Gamma^\S\,\gamma\,\gamma^D\ra(\alpha:A^\A\,\gamma)\ra A^\D\,\gamma^D\,\alpha\ra\Set)
  \end{alignat*}
  Substitutions (and terms) are given by (dependent) functions which
  respect the predicates $^\D$ and relations $^\S$.
  \begin{alignat*}{10}
    & (\sigma:\Sub\,\Gamma\,\Delta) && := \,\, && (\sigma^\A:\Gamma^\A\ra\Delta^\A) \times (\sigma^\D:\Gamma^\D\,\gamma\ra\Delta^\D\,(\sigma^\A\,\gamma)) \times \\
    & && && \rlap{$(\sigma^\S:\Gamma^\S\,\gamma\,\gamma^D\ra\Delta^\S\,(\sigma^A\,\gamma)\,(\sigma^\D\,\gamma^D))$} \\
    & (t:\Tm\,\Gamma\,A) && := && (t^\A:(\gamma:\Gamma^\A)\ra A^\A\,\gamma) \times (t^\D:(\gamma^D:\Gamma^\D\,\gamma)\ra A^\D\,\gamma^D\,(t^\A\,\gamma)) \times \\
    & && && (t^\S:(\gamma^S:\Gamma^\S\,\gamma\,\gamma^D)\ra A^\S\,(t^A\,\gamma)\,(t^\D\,\gamma^D))
  \end{alignat*}

>>>>

\begin{example}[Example signatures]\label{ex:signatures}
  The signature for natural numbers is given by the function
  \[
    nat := \lambda M.(\boldsymbol{\cdot}^M\,\ext^M\,\U^M\,\ext^M\,\El^M\,0^M\,\ext^M\,1^M \Ra^M \El^M\,1^M).
  \]
  For readability we write the same signature without the $\lambda$
  and the superscripts below on the left and we also write an informal
  version with variable names on the right for comparison.
  \[
  \boldsymbol{\cdot}\,\ext\,\U\,\ext\,\El\,0\,\ext\,1 \Ra \El\,1 \hspace{5.9em}
  \boldsymbol{\cdot}\,\ext\,N:\U\,\ext\,z : \El\,N\,\ext\,s : N \Ra \El\,N
  \]
  This signature is given by a context of length three. We start with
  the empty context $\boldsymbol{\cdot}$, then we declare a sort by
  $\U$, then we declare an operator producing an element of $\El\,0$
  where $0$ is a de Bruijn index referring to the sort $\U$. Finally
  we declare an operator which takes as input an element of the sort
  $1$ and produces an element of the same sort. Note the assymetry of
  the function type $\Ra$: on the left there needs to be an element of
  $\U$, while on the right any type can appear (including another
  function type). This ensures strict positivity of the operators.

@@

  $\Con$, $\Ty$, $\Sub$, and so on, one per row. Each such components

@@

\begin{definition}[A model of ETT supports IITs]\label{sec:hasiits}
  A model of ETT supports IITs if for any signature signature
  $f:(M:\SignAlg)\ra\Con^M$ we have a $\con_f: (f\,\ADS)^\A$ and an
  $\elim_f : (\gamma^D:(f\,\ADS)^\D\,\con_f)\ra (f\,\ADS)^\S\,\con_f\,\gamma^D$.
\end{definition}
In other words, for any signature, we have an algebra called $\con$
(constructors) and for any displayed algebra over the constructors, we
have a section.

One can check that \Cref{sec:hasiits} gives the right
notion of constructors and elimination principle for the signatures in
\Cref{ex:signatures}.
\begin{example}[An IIT supports natural numbers]
  For the signature $nat$ of natural numbers in \Cref{ex:signatures},
  the type of $\con_{nat}$ is
  \begin{alignat*}{5}
    & ({nat}\,\ADS)^\A = \\
    & (\boldsymbol{\cdot}^\ADS\,\ext^\ADS\,\U^\ADS\,\ext^\ADS\,\El^\ADS\,0^\ADS\,\ext^\ADS\,1^\ADS \Ra^\ADS \El^\ADS\,1^\ADS)^\A = \\
    & \Big(\big((\boldsymbol{\cdot}\,\ext\,\U)\,\ext\,\El\,(\pi_2\,\id)\big)\,\ext\,(\pi_2\,(\pi_1\,\id)) \Ra \El\,(\pi_2\,(\pi_1\,\id))\Big)^\A = \\
    & \Big(\gamma'':\big(\gamma':((\gamma:\top)\times\Set)\big)\times(\proj_2\,\gamma')\Big)\times\big(\proj_2\,(\proj_1\,\gamma'')\ra\proj_2\,(\proj_1\,\gamma'')\big),
  \end{alignat*}
  which is the left-associated $\Sigma$ type isomorphic to its
  right-associated counterpart
  \[
    (N:\Set)\times \big(N \times (N\ra N)\big).
  \]
  Pretending that we always had the right-associated variant, the type
  of $\elim_{nat}$ computes as follows.
  \begin{alignat*}{5}
    & (\gamma^D:\,\, && ({nat}\,\ADS)^\D\,\con_{nat})\ra ({nat}\,\ADS)^\S\,\con_{nat}\,\gamma^D = \\
    & (\gamma^D:\,\, && (\boldsymbol{\cdot}^\ADS\,\ext^\ADS\,\U^\ADS\,\ext^\ADS\,\El^\ADS\,0^\ADS\,\ext^\ADS\,1^\ADS \Ra^\ADS \El^\ADS\,1^\ADS)^\D\,\con_{nat})\ra \\
    & && (\boldsymbol{\cdot}^\ADS\,\ext^\ADS\,\U^\ADS\,\ext^\ADS\,\El^\ADS\,0^\ADS\,\ext^\ADS\,1^\ADS \Ra^\ADS \El^\ADS\,1^\ADS)^\S\,\con_{nat}\,\gamma^D = \\
    & (\gamma^D:\,\, && (\boldsymbol{\cdot}^\ADS\,\ext^\ADS\,\U^\ADS\,\ext^\ADS\,\El^\ADS\,0^\ADS\,\ext^\ADS\,1^\ADS \Ra^\ADS \El^\ADS\,1^\ADS)^\D\,\con_{nat})\ra \\
    & && (\boldsymbol{\cdot}^\ADS\,\ext^\ADS\,\U^\ADS\,\ext^\ADS\,\El^\ADS\,0^\ADS\,\ext^\ADS\,1^\ADS \Ra^\ADS \El^\ADS\,1^\ADS)^\S\,\con_{nat}\,\gamma^D = \\
  \end{alignat*}
\end{example}
